---
title: "AudioConversation Mel"
format: html
editor: source
---

Converts audio file into Mel Frequency Cepstrum Coefficients. 
Time is split into 0.10 second long chunks, with a 0.05 second overlap.


```{r}
library(tuneR)
library(seewave)
library(tidyr)
library(dplyr)
library(readxl)
```





## Reading in files

```{r}
audio_dataAW <-readWave(here::here("TrainingData", "AW", "671658014.180928183606-AW.wav"),
                        from = 0, to = 0.5, units = "hours")
```

```{r}
audio_dataAS <-readWave(here::here("TrainingData", "AS", "671658014.180929033558-AS.wav"),
                        from = 0, to = 0.5, units = "hours")
```

```{r}
audio_dataSS <-readWave(here::here("TrainingData", "SS", "671658014.181007213415-SS.wav"),
                        from = 0, to = 0.5, units = "hours")

```

## Extracting Raw Audio Samples

```{r}
raw_samplesAW <- audio_dataAW@left #mono data has identical left and right values
```

```{r}
raw_samplesAS <- audio_dataAS@left #mono data has identical left and right values
```

```{r}
raw_samplesSS <- audio_dataSS@left #mono data has identical left and right values
```

## Computing Mel Coefficients

```{r}
melcoeffsAW = melfcc(audio_dataAW, wintime = 0.1, hoptime = 0.05, numcep = 12, bwidth = 1)


melcoeffsAWlowerhz = melfcc(audio_dataAW, wintime = 0.1, hoptime = 0.05, 
                            numcep = 12, bwidth = 1, minfreq = 0, maxfreq = 2000)
```

```{r}
#melcoeffsAS = melfcc(audio_dataAS, wintime = 0.1, hoptime = 0.1, numcep = 12, bwidth = 1)
```

```{r}
#melcoeffsSS = melfcc(audio_dataSS, wintime = 0.1, hoptime = 0.1, numcep = 12, bwidth = 1)

```

## Adding Time Steps to Matrix

```{r}
#not sure if these are the correct time step bounds, investigate further

#17999 is for 30 minutes
timestepsAW = as.data.frame(cbind(melcoeffsAW, 
              "Time_Start" = seq(from = 0, to = 1799.9, by = 0.05),
              "Time_End" = seq(from = .1, to = 1799.95, by = 0.05)))


#17999 is for 30 minutes
timestepsAWlowerhz = as.data.frame(cbind(melcoeffsAWlowerhz, 
              "Time_Start" = seq(from = 0, to = 1799.9, by = 0.05),
              "Time_End" = seq(from = .1, to = 1799.95, by = 0.05)))

```

```{r}
#timestepsAS = as.data.frame(cbind(melcoeffsAS, 
#              "Time_Start" = seq(from = 0, to = 1799.8, length.out = 17999),
#              "Time_End" = seq(from = .1, to = 1799.9, length.out = 17999)))
```

```{r}
timestepsSS = as.data.frame(cbind(melcoeffsSS, 
              "Time_Start" = seq(from = 0, to = 1799.8, length.out = 17999),
              "Time_End" = seq(from = .1, to = 1799.9, length.out = 17999)))

```

```{r}
#write.csv(timesteps, "timestepsAW.csv")
```

## Reading in a Training Data Set

```{r}
annotatedAW = read_excel(here::here("TrainingData", "AW", "AWhb.xlsx"))

annotatedAWlowerhz = read_excel(here::here("TrainingData", "AW", "AWhb.xlsx"))

```

```{r}

annotatedAS = read_excel(here::here("TrainingData", "AS", "AScopy.xlsx"))
```

```{r}

annotatedSS = read_excel(here::here("TrainingData", "SS", "SShb.xlsx"))
```

## Cleaning Training Data Set

```{r}
annotatedAW = annotatedAW %>% rename(Begin_Time = 'Begin Time (s)',
                     End_Time = 'End Time (s)')

annotatedAWlowerhz = annotatedAWlowerhz %>% rename(Begin_Time = 'Begin Time (s)',
                     End_Time = 'End Time (s)')
```

```{r}
#annotatedAS = annotatedAS %>% rename(Begin_Time = 'Begin Time (s)',
#                     End_Time = 'End Time (s)')
```

```{r}
annotatedSS = annotatedSS %>% rename(Begin_Time = 'Begin Time (s)',
                     End_Time = 'End Time (s)')
```


## Extracting Ranges of Calls

```{r}

whalerangesAW = apply(cbind(annotatedAW$Begin_Time, annotatedAW$End_Time)
, 1, list)


whalerangesAWlowerhz = apply(cbind(annotatedAWlowerhz$Begin_Time, annotatedAWlowerhz$End_Time)
, 1, list)
```

```{r}
#whalerangesAS = apply(cbind(annotatedAS$Begin_Time, annotatedAS$End_Time)
#, 1, list)
```

```{r}
whalerangesSS = apply(cbind(annotatedSS$Begin_Time, annotatedSS$End_Time)
, 1, list)


#whaleranges[[1]][[1]]

#whaleranges
```

## Check if Number is within the Range of any list of Numbers (AI Code - modified)

```{r}
#' Checks if a number falls within any of a list of specified ranges.
#'
#' @param number The numeric value to check.
#' @param ranges A list of numeric vectors, where each vector represents a range
#'   with two elements: `c(start, end)`.
#' @return A logical value: `TRUE` if the number is within any range, `FALSE` otherwise.
#' @examples
#' check_in_ranges(5, list(c(1, 10), c(15, 20))) # Returns TRUE
#' check_in_ranges(12, list(c(1, 10), c(15, 20))) # Returns FALSE
#' check_in_ranges(18, list(c(1, 10), c(15, 20))) # Returns TRUE
check_in_ranges <- function(number, ranges, timesteplength = 0.1) {
  # Initialize a flag to track if the number is found within any range
  found <- FALSE

  # Iterate through each range in the list
  for (r in 1:length(ranges)) {
    #print(r)
    #print(ranges[[r]][[1]][2])
    # Check if the number is within the current range (inclusive)
    if (number <= ranges[[r]][[1]][2] && 
        number + timesteplength >= ranges[[r]][[1]][1]) {
      found <- TRUE
      break # Exit the loop as soon as a match is found
    }
  }

  return(c(number, found))
}
```

```{r}
#testing function
#lapply(c(1,2), check_in_ranges, ranges = whalerangesAW)
```

## Apply Ranges Function to Timesteps

```{r}

overlapstartAW = lapply(timestepsAW$Time_Start, check_in_ranges, ranges = whalerangesAW)

overlapstartAWlowerhz = lapply(timestepsAWlowerhz$Time_Start, check_in_ranges, ranges = whalerangesAWlowerhz)

```

```{r}
#overlapstartAS = lapply(timestepsAS$Time_Start, check_in_ranges, ranges = whalerangesAS)
```

```{r}
overlapstartSS = lapply(timestepsSS$Time_Start, check_in_ranges, ranges = whalerangesSS)

```

## Merging Output of Function With Mel Coefficient Datasets

```{r}

#data frame stating whether start of time window is contained in a hb whale call
startmatcheddfAW = as.data.frame(t(as.data.frame(overlapstartAW))) %>%
  rename(Time_Start = V1, Humpback = V2)


annotatedfullAW = full_join(timestepsAW, startmatcheddfAW, 
                          by = join_by(Time_Start == Time_Start))



startmatcheddfAWlowerhz = as.data.frame(t(as.data.frame(overlapstartAWlowerhz))) %>%
  rename(Time_Start = V1, Humpback = V2)


annotatedfullAWlowerhz = full_join(timestepsAWlowerhz, startmatcheddfAWlowerhz, 
                          by = join_by(Time_Start == Time_Start))
```

```{r}
#startmatcheddfAS = as.data.frame(t(as.data.frame(overlapstartAS))) %>%
#  rename(Time_Start = V1, Humpback = V2)


#annotatedfullAS = full_join(timestepsAS, startmatcheddfAS, 
#                          by = join_by(Time_Start == Time_Start))

```

```{r}
startmatcheddfSS = as.data.frame(t(as.data.frame(overlapstartSS))) %>%
  rename(Time_Start = V1, Humpback = V2)


annotatedfullSS = full_join(timestepsSS, startmatcheddfSS, 
                          by = join_by(Time_Start == Time_Start))


```

## Adding Instance Numbers

```{r}

add_instances =  function(dataframe){
  
  
  #
  dataframe$instances = rep(0, times = nrow(dataframe))
  dataframe$initial = rep(0, times = nrow(dataframe))
  
  #recording successive positive/negative observations
  dataframe$streak = rep(0, times = nrow(dataframe))
  
  instances = 0
  
  for (i in 1:nrow(dataframe)) {
    # Access the current row using df[i]
    current = dataframe$Humpback[i]
    
    
    #if on the first row, assume the previous was 0
    if (i == 1) {
      previous = 0
    }
    
    #if not on first row, set previous value to use as comparison
    if (i > 1) {
      previous = dataframe$Humpback[i - 1]
      
    }
    
    #if previous has no whale call and current has whale call
    if (previous == 0 & current == 1){
      #create new instance since it is a new whale call
      instances = instances + 1
      dataframe$instances[i] = instances
      
      #indiciate that this is the initilization of a new whale call
      dataframe$initial[i] = 1
      
    }
    
    #if previous has a whale call and current also has a whale call
    if (previous == 1 & current == 1){
      dataframe$instances[i] = instances
    }
  }
  
  result =  dataframe
}
```

```{r}
annotatedfullAW = add_instances(annotatedfullAW)


annotatedfullAWlowerhz = add_instances(annotatedfullAWlowerhz)

```

```{r}
annotatedfullSS = add_instances(annotatedfullSS)
```

```{r}
write.csv(annotatedfullAW, "annotatedmelAWoverlap.csv")

write.csv(annotatedfullAWlowerhz, "annotatedmelAWlowerhzoverlap.csv")

```

```{r}
#write.csv(annotatedfullAS, "annotatedmelAS.csv")
```

```{r}
write.csv(annotatedfullSS, "annotatedmelSS.csv")

```
