---
title: "CreatingContextWindows2"
format: html
editor: source
---

## Context Windows

The context window length, corresponds to a 3.9124-s long spectrogram image. A context window of 3.9124-s length is large enough to contain an entire unit but not large enough to contain two units (on average).

## Loading Libraries

```{r}
library(tuneR)
library(seewave)
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)

library(tidyverse)
library(tidymodels)
library(kknn)
library(discrim)
```


```{r}
source("ContextWindowFunctions.R")

```


## Loading In Data

```{r}
annotatedMelAW = read.csv(here::here("annotatedmelAWoverlap.csv"))
# first 30 minutes is in this data set

#annotatedMelAW %>% arrange(desc(Time_End))
```







### Creating Training Data

```{r}
trainingdata = add_observations(whaledf = annotatedMelAW,
                                start_time = 0,
                                end_time = 200,
                                num_time_steps=40, 
                                numcep = 12,
                                step_size = 0.05)


table(trainingdata$Humpback)
```

### Creating Testing Data


```{r}

testingdata = add_observations(whaledf = annotatedMelAW,
                                start_time = 200,
                                end_time = 400,
                                num_time_steps=40, 
                                numcep = 12,
                                step_size = 0.05)



table(testingdata$Humpback)

```


### Creating Cross-Validation Folds

```{r}
set.seed(326)
training_cvs = vfold_cv(trainingdata, v = 10)

```

## LDA

### Creating LDA Model

```{r}
#creating model
lda_mod1 = discrim_linear() %>%
  set_engine("MASS") %>%
  set_mode("classification")

#creating recipe
ldarecipe_1 = recipe(Humpback ~ ., data = trainingdata) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
ldaworkflow_1 = workflow() %>% 
  add_recipe(ldarecipe_1) %>% 
  add_model(lda_mod1)

#fitting model
ldafit_1 = ldaworkflow_1 %>% fit(trainingdata) 

#collinearity warning sometimes

#confusion matrix
trainingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, trainingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = ldapreds)
```


```{r}
trainingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, trainingdata)$.pred_class
  ) %>% select(Humpback, ldapreds, Time_Start, Time_End)


#ldaworkflow_1 %>% fit_resamples(training_cvs)


ldafitcv_1 = ldaworkflow_1 %>% fit_resamples(training_cvs,
                metrics = metric_set(accuracy, roc_auc, precision, specificity, recall)
                )
collect_metrics(ldafitcv_1)


2236/(2336 + 60)
```

### Predicting on Testing Data

```{r}

#confusion matrix
testingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, testingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = ldapreds)



testingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, testingdata)$.pred_class
  ) %>% 
  select(Humpback, ldapreds, Time_Start, Time_End)



testingdata %>%
  mutate(
    ldapreds = as.factor(predict(ldafit_1, testingdata)$.pred_class)
  ) %>% recall(truth = Humpback, estimate =  ldapreds, event_level = 'second')
```

### Computing Metrics on Validation Data Set

```{r}

```



### Create Output For Raven Visualization




```{r}
testingoutput = testingdata %>%
  mutate(ldapreds = predict(ldafit_1, testingdata)$.pred_class) %>% 
  dplyr::select(Time_Start, Time_End, Humpback, ldapreds)




#creating dummy low and high frequencies for now

testingoutput = testingoutput %>% mutate(low = rep(0, times = nrow(testingoutput)),
                         high = rep(1000, times = nrow(testingoutput)))



testingoutput = testingoutput %>% filter(ldapreds == 1) %>% 
  rename("Begin Time (s)" = Time_Start, "End Time (s)" = Time_End, 
         "Low Freq (Hz)" = low, "High Freq (Hz)" = high) %>% 
  dplyr::select(c("Begin Time (s)", "End Time (s)", 
                  "Low Freq (Hz)", "High Freq (Hz)"))



#create output 



write.table(testingoutput, "fulloverlapldapreds.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

## KNN

### Creating KNN Model

```{r}
#creating model
knn_mod1 = nearest_neighbor(neighbors = 100) %>%
  set_engine("kknn") %>%
  set_mode("classification")

#creating recipe
knnrecipe_1 = recipe(Humpback ~ ., 
                     data = trainingdata) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
knnworkflow_1 = workflow() %>% 
  add_recipe(knnrecipe_1) %>% 
  add_model(knn_mod1)

#fitting model
knnfit_1 = knnworkflow_1 %>% fit(trainingdata) 


#confusion matrix
trainingdata %>%
  mutate(
    knnpreds = predict(knnfit_1, trainingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = knnpreds)




```


### KNN Predicting on Testing

```{r}

#confusion matrix
testingdata %>%
  mutate(
    knnpreds = predict(knnfit_1, testingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = knnpreds)

testingdata %>%
  mutate(
    knnpreds = as.factor(predict(knnfit_1, testingdata)$.pred_class)
  ) %>% recall(truth = Humpback, estimate =  knnpreds, event_level = 'second')


```

## Decision Tree



### Creating Model

```{r}
#model
decision_tree_mod1 = decision_tree() %>%
  set_engine("rpart") %>%
  set_mode("classification")

#recipe
treerecipe_1 = recipe(Humpback ~ ., 
                     data = trainingdata) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())

#workflow
tree_wflow = workflow() %>%
  add_model(decision_tree_mod1) %>% 
  add_recipe(treerecipe_1)


#fitting model
tree_fit_1 = tree_wflow %>% fit(trainingdata) 


#confusion matrix
trainingdata %>%
  mutate(
    treepreds = predict(tree_fit_1, trainingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = treepreds)
```

### DT Predicting on Testing



```{r}
#confusion matrix
testingdata %>%
  mutate(
    treepreds = predict(tree_fit_1, testingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = treepreds)

testingdata %>%
  mutate(
    treepreds = as.factor(predict(tree_fit_1, testingdata)$.pred_class)
  ) %>% recall(truth = Humpback, estimate =  treepreds, event_level = 'second')


```




