---
title: "Creating Context Windows"
format: html
editor: source
---


This has different functions for creating positive and negative observations,
which is not necessary. Use 'Creating ContextWindowsNEW.qmd' instead.






## Loading Libraries

```{r}
library(tuneR)
library(seewave)
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)

library(tidyverse)
library(tidymodels)
library(kknn)
library(discrim)
```

## Loading In Data

```{r}
annotatedMelAW = read.csv(here::here("annotatedmelAWoverlap.csv"))
# first 30 minutes is in this data set

#annotatedMelAW %>% arrange(desc(Time_End))
```




## Examining Call Durations

```{r}
#examining the relationship between call length and when call occurred

annotatedAW = read_excel(here::here("TrainingData", "AW", "AWhb.xlsx"))


annotatedAW %>% ggplot(aes(x = `Begin Time (s)`, y = callduration)) +
  geom_point() + labs(title = "Call Length Over Time") + ylim(0,5)


#histogram of cull durations
median(annotatedAW$callduration)

annotatedAW %>% ggplot(aes(x = callduration)) +
  geom_histogram(bins = 200) + 
  labs(title = "Distribution of Whale Call Lengths") +
  geom_vline(xintercept = median(annotatedAW$callduration), col = "red")
```



## Testing Making First Context Window

```{r}
#probably best to use a function

newdf = annotatedMelAW %>% filter(Time_End <= 4)

as.vector(newdf[, 1:13])

#takes in a specified number of 
test_function = function(i){

  testdf = data.frame(matrix(nrow = 1, ncol = 10))
  print(testdf)
  colnames(testdf) = paste0("V1_",i)
  print(testdf)
}


test_function(5)
```

## Creating Empty Data Frame


```{r}

#numcep is the number of mel coefficients used
#num_time_steps is the number of time steps recorded in an observation

#create_observations takes in the above parameters,
  #outputs a data frame with num_time_steps*numcep columns, 
  #and 'Time_Start', 'Time_End' and Humpback (0/1) columns

create_observations = function(numcep, num_time_steps){
  #testdf = data.frame(nrow = num_time_steps*numcep)
  #print(testdf)
  colnamelist = list()
  for (i in 1:numcep){
    for (j in 1:num_time_steps){
      #print(i)
      #print(j)
      #testdf = cbind(testdf, paste0("V1_",i))
      colnamelist = append(colnamelist, paste0("V", i, "_", j))
    }
  }
  #adding Time_Start and Time_End variables to end of each row
  colnamelist = append(colnamelist, "Time_Start")
  colnamelist = append(colnamelist, "Time_End")
  colnamelist = append(colnamelist, "Humpback")


  dummydf = data.frame(matrix(nrow = 1, ncol = num_time_steps*numcep + 3))
  
  colnames(dummydf) = colnamelist
  
  
  
  
  
  return(dummydf)
  
}
#emptydf = create_observations(numcep = 8, num_time_steps = 40)

#emptydf
```



## Creating Observations


```{r}

#function for specifically overlapped time steps (0.5 second increases, 0.5 second overlap)


# dummydf is the empty df created in the above 'create_observations' function
# whaledf is the full mel coefficient data frame
# num_time_steps is the number of time steps 
# numcep is the number of cepstrum coefficients



add_observations = function(whaledf, num_time_steps, numcep){
  
  
  dummydf = create_observations(numcep, num_time_steps)

  #getting initial times
  #filtering on initial times
  initials = whaledf %>% pull(Time_Start)
  
  # for each initial row in which an observation is desired
  for (i in initials){
    
    
    
    #i represents the start time of an initial observation (context window)
    
    #create observation by looping over the number of time steps
    
    #if statement to avoid including observations at end of the file
    if (nrow(whaledf %>% 
             filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps)) == 
        num_time_steps)
      {
        # determining whether the window has any whale call
        if (nrow(whaledf %>% 
             filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps) %>%
             filter(Humpback == 1)) >= 1){
          
        whale = 1
        } else {
        whale = 0
        }
   
          
      coeffs = whaledf %>% 
        filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps) %>% 
        dplyr::select(V1:paste0("V", numcep))
    
    
      #c(as.matrix()) creates a vector out of a data.frame, going down columns
      dummydf = rbind(dummydf, 
                      c(as.matrix(coeffs), i, i + 0.05*num_time_steps, whale))
      }
    
    
    #need to create conditionals for end of file
  }
  return(dummydf)
}


#positives = 
  
  
qwerty = add_observations(whaledf = annotatedMelAW %>% 
                   filter(0 <= Time_Start & Time_Start < 20),num_time_steps=40, 
                          numcep = 12) %>% drop_na()


table(qwerty$Humpback)


```




## Creating Positive Observations


```{r}

#function for specifically overlapped time steps (0.5 second increases, 0.5 second overlap)


# dummydf is the empty df created in the above 'create_observations' function
# whaledf is the full mel coefficient data frame
# num_time_steps is the number of time steps 
# numcep is the number of cepstrum coefficients


# positive observations are created by initializing context windows 
# on every time window that contains any whale call
# the window lasts for the specified time length
add_positive_observations = function(whaledf, num_time_steps, numcep){
  
  
  dummydf = create_observations(numcep, num_time_steps)

  #getting initial times
  #filtering on initial times
  initials = whaledf %>% filter(initial == 1) %>% pull(Time_Start)
  
  # for each initial row in which an observation is desired
  for (i in initials){
    #i represents the start time of an initial observation (context window)
    
    #create observation by looping over the number of time steps
    
    #if statement to avoid including observations at end of the file
    if (nrow(whaledf %>% 
             filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps)) == 
        num_time_steps)
      {
    
   
          
      coeffs = whaledf %>% 
        filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps) %>% 
        dplyr::select(V1:paste0("V", numcep))
    
    
      #c(as.matrix()) creates a vector out of a data.frame, going down columns
      dummydf = rbind(dummydf, c(as.matrix(coeffs), i, i + 0.05*num_time_steps, 1))
      }
    
    
    #need to create conditionals for end of file
  }
  return(dummydf)
}


#positives = 
  
  
add_positive_observations(whaledf = annotatedMelAW %>% filter(0 <= Time_Start & Time_Start < 100), num_time_steps = 40, 
                          numcep = 1) %>% drop_na()





```



## Creating Negative Observations


```{r}
add_negative_observations = function(whaledf, num_time_steps, numcep){
  
  
  dummydf = create_observations(numcep, num_time_steps)
  
  #if an initial negative and no new whale call for the entire context window
  
  
  #getting initial times
  initials = whaledf %>% filter(Humpback == 0) %>% pull(Time_Start)
  
  # for each initial row in which an observation is desired
  for (i in initials){
    #i represents the start time of an initial observation
    
    #create observation by looping over the number of time steps
  
 
    # making sure there are only 0's and the proper amount of time steps
    if (nrow(whaledf %>% 
      filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps) %>% 
        filter(Humpback == 0)) == num_time_steps) {

    
      coeffs = whaledf %>% 
        filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps) %>% 
        dplyr::select(V1:paste0("V", numcep))
      
    
      #c(as.matrix()) creates a vector out of a data.frame, going down columns
      dummydf = 
        rbind(dummydf, c(as.matrix(coeffs), i, i + 0.05*num_time_steps, 0))
      }
    
  }
  return(dummydf)
}


#negatives = add_negative_observations(
#  annotatedMelAW %>% filter(Time_Start < 1800), 
#                          num_time_steps = 40, 
#                          numcep = 8) %>% drop_na()

# overfitting on negative due to frequent overlaps?

```



## Creating Non-Overlapping Context Windows


```{r}
# very similar to the above function

# add conditional statement making sure start of context window does not overlap with previous context window




add_nonoverlaping_negative_observations = function(whaledf, num_time_steps, numcep)
    {
  previoustime = 0
  
  dummydf = create_observations(numcep, num_time_steps)
  
  #if an initial negative and no new whale call for the entire context window
  
  
  #getting initial times
  initials = whaledf %>% filter(Humpback == 0) %>% pull(Time_Start)
  
  # for each initial row in which an observation is desired
  for (i in initials){
    #i represents the start time of an initial observation
    
    #create observation by looping over the number of time steps
    
    
 
    # making sure there are only 0's and the proper amount of time steps
    if (nrow(whaledf %>% 
      filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps) %>% 
        filter(Humpback == 0)) == num_time_steps &
      previoustime + 0.05*num_time_steps <= i)
      {

    
      coeffs = whaledf %>% 
        filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps) %>% 
        dplyr::select(V1:paste0("V", numcep))
      
    
      #c(as.matrix()) creates a vector out of a data.frame, going down columns
      dummydf = 
        rbind(dummydf, c(as.matrix(coeffs), i, i + 0.05*num_time_steps, 0))
      
      previoustime = i
      
      
      
      
      }
    
    }
  return(dummydf)
  }


#negatives = 

#add_nonoverlaping_negative_observations(
#  annotatedMelAW %>% filter(Time_Start < 1800), 
#                          num_time_steps = 40, 
#                          numcep = 1) %>% drop_na()




```



## Cross Validation


### Combining Positive and Negative Observations for Cross Validation

### Creating Training Data

```{r}
trainingdata = rbind(add_positive_observations(
  whaledf = annotatedMelAW %>% filter(Time_Start < 1000),
                          num_time_steps = 40, 
                          numcep = 12) %>% drop_na(), 
                     add_negative_observations(
  whaledf = annotatedMelAW %>% filter(Time_Start < 1000), 
                          num_time_steps = 40, 
                          numcep = 12) %>% drop_na())

trainingdata = trainingdata %>% mutate(Humpback = as.factor(Humpback))

#trainingdata %>% filter(Humpback ==1)

table(trainingdata$Humpback)
```

### Creating Testing Data


```{r}
add_testing_observations = function(whaledf, num_time_steps, numcep){
  
  
  dummydf = create_observations(numcep, num_time_steps)
  
  #if an initial negative and no new whale call for the entire context window
  
  
  #getting initial times
  initials = whaledf %>% pull(Time_Start)
  
  # for each initial row in which an observation is desired
  for (i in initials){
    #create observation by looping over the number of time steps
    
    # determining whether a humpback is present at this time step
    presence = whaledf %>% filter(Time_Start == i) %>% pull(Humpback)
 
    # making sure there are the proper amount of time steps
    if (nrow(whaledf %>% 
      filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps)) 
      == num_time_steps) {

    
      coeffs = whaledf %>% 
        filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps) %>% 
        dplyr::select(V1:paste0("V", numcep))
      
    
      #c(as.matrix()) creates a vector out of a data.frame, going down columns
      dummydf = 
        rbind(dummydf, c(as.matrix(coeffs), i, i + 0.05*num_time_steps, presence))
      }
    
  }
  return(dummydf)
}


#annotatedMelAW %>% filter(Time_Start == 1788) %>% pull(Humpback)
```

```{r}

testingdata = add_testing_observations(
  annotatedMelAW %>% filter(1700 <= Time_Start & Time_Start < 1800), 
                          num_time_steps = 40, 
                          numcep = 12) %>% drop_na()


testingdata = testingdata %>% mutate(Humpback = as.factor(Humpback))



```


```{r}
#history of trials
  #1. first 1800 seconds, 12 coeffs, 40 time steps, huge overlap in negatives
    # recall: 0.9876399
  
  #2.
    #training on first 1000 seconds
    #testing on remaining 8000 seconds

```



### Creating Cross-Validation Folds

```{r}
set.seed(130)
training_cvs = vfold_cv(trainingdata, v = 10)

```

### Creating LDA Model

```{r}
#creating model
lda_mod1 = discrim_linear() %>%
  set_engine("MASS") %>%
  set_mode("classification")

#creating recipe
ldarecipe_1 = recipe(Humpback ~ ., data = trainingdata) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
ldaworkflow_1 = workflow() %>% 
  add_recipe(ldarecipe_1) %>% 
  add_model(lda_mod1)

#fitting model
ldafit_1 = ldaworkflow_1 %>% fit(trainingdata) 


#confusion matrix
trainingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, trainingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = ldapreds)
```


```{r}
trainingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, trainingdata)$.pred_class
  )


#ldaworkflow_1 %>% fit_resamples(training_cvs)


ldafitcv_1 = ldaworkflow_1 %>% fit_resamples(training_cvs,
                metrics = metric_set(accuracy, roc_auc, precision, specificity, recall)
                )
collect_metrics(ldafitcv_1)


2236/(2336 + 60)
```

### LDA Predicting on Testing Data

```{r}
#fitting model
ldafit_2 = ldaworkflow_1 %>% fit(trainingdata)


#confusion matrix
testingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, testingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = ldapreds)

predict(ldafit_1, trainingdata)$.pred_class

```

```{r}
trainingdata
```






### LDA Computing Metrics on Validation Data Set

```{r}

```



### LDA Create Output For Raven Visualization


```{r}
testingoutput = testingdata %>%
  mutate(ldapreds = predict(ldafit_1, testingdata)$.pred_class) %>% 
  dplyr::select(Time_Start, Time_End, Humpback, ldapreds)




#creating dummy low and high frequencies for now

testingoutput = testingoutput %>% mutate(low = rep(0, times = nrow(testingoutput)),
                         high = rep(1000, times = nrow(testingoutput)))



testingoutput = testingoutput %>% filter(ldapreds == 1) %>% 
  rename("Begin Time (s)" = Time_Start, "End Time (s)" = Time_End, 
         "Low Freq (Hz)" = low, "High Freq (Hz)" = high) %>% 
  dplyr::select(c("Begin Time (s)", "End Time (s)", 
                  "Low Freq (Hz)", "High Freq (Hz)"))



#create output 



write.table(testingoutput, "overlapldapreds.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```


### KNN

```{r}
#creating model
knn_mod1 = nearest_neighbor(neighbors = 100) %>%
  set_engine("kknn") %>%
  set_mode("classification")

#creating recipe
knnrecipe_1 = recipe(Humpback ~ ., 
                     data = trainingdata) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
knnworkflow_1 = workflow() %>% 
  add_recipe(knnrecipe_1) %>% 
  add_model(knn_mod1)

#fitting model
knnfit_1 = knnworkflow_1 %>% fit(trainingdata) 


#confusion matrix
trainingdata %>%
  mutate(
    knnpreds = predict(knnfit_1, trainingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = knnpreds)




```


### KNN Predicting on Testing

```{r}

#confusion matrix
testingdata %>%
  mutate(
    knnpreds = predict(knnfit_1, testingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = knnpreds)

#predict(knnfit_1, trainingdata)$.pred_class




