---
title: "CreatingContextWindows2"
format: html
editor: source
---

## Context Windows

The context window length, corresponds to a 3.9124-s long spectrogram image. A context window of 3.9124-s length is large enough to contain an entire unit but not large enough to contain two units (on average).

## Loading Libraries

```{r}
library(tuneR)
library(seewave)
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)

library(tidyverse)
library(tidymodels)
library(kknn)
library(discrim)
```

## Loading In Data

```{r}
annotatedMelAW = read.csv(here::here("annotatedmelAWoverlap.csv"))
# first 30 minutes is in this data set

#annotatedMelAW %>% arrange(desc(Time_End))
```







## Creating Empty Data Frame


```{r}

#numcep is the number of mel coefficients used
#num_time_steps is the number of time steps recorded in an observation

#create_observations takes in the above parameters,
  #outputs a data frame with num_time_steps*numcep columns, 
  #and 'Time_Start', 'Time_End' and Humpback (0/1) columns

create_observations = function(numcep, num_time_steps){
  #testdf = data.frame(nrow = num_time_steps*numcep)
  #print(testdf)
  colnamelist = list()
  for (i in 1:numcep){
    for (j in 1:num_time_steps){
      #print(i)
      #print(j)
      #testdf = cbind(testdf, paste0("V1_",i))
      colnamelist = append(colnamelist, paste0("V", i, "_", j))
    }
  }
  #adding Time_Start and Time_End variables to end of each row
  colnamelist = append(colnamelist, "Time_Start")
  colnamelist = append(colnamelist, "Time_End")
  colnamelist = append(colnamelist, "Humpback")


  dummydf = data.frame(matrix(nrow = 1, ncol = num_time_steps*numcep + 3))
  
  colnames(dummydf) = colnamelist
  
  
  
  
  
  return(dummydf)
  
}
#emptydf = create_observations(numcep = 8, num_time_steps = 40)

#emptydf
```



## Creating Observations


```{r}

#function for specifically overlapped time steps (0.5 second increases, 0.5 second overlap)


# dummydf is the empty df created in the above 'create_observations' function
# whaledf is the full mel coefficient data frame
# num_time_steps is the number of time steps 
# numcep is the number of cepstrum coefficients



add_observations = function(whaledf, num_time_steps, numcep){
  
  
  dummydf = create_observations(numcep, num_time_steps)

  #getting initial times
  #filtering on initial times
  initials = whaledf %>% pull(Time_Start)
  
  # for each initial row in which an observation is desired
  for (i in initials){
    
    
    
    #i represents the start time of an initial observation (context window)
    
    #create observation by looping over the number of time steps
    
    #if statement to avoid including observations at end of the file
    if (nrow(whaledf %>% 
             filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps)) == 
        num_time_steps)
      {
        # determining whether the window has any whale call
        if (nrow(whaledf %>% 
             filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps) %>%
             filter(Humpback == 1)) >= 1){
          
        whale = 1
        } else {
        whale = 0
        }
   
          
      coeffs = whaledf %>% 
        filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps) %>% 
        dplyr::select(V1:paste0("V", numcep))
    
    
      #c(as.matrix()) creates a vector out of a data.frame, going down columns
      dummydf = rbind(dummydf, 
                      c(as.matrix(coeffs), i, i + 0.05*num_time_steps, whale))
      }
    
    
    #need to create conditionals for end of file
  }
  return(dummydf)
}


#positives = 
  
  



table((add_observations(whaledf = annotatedMelAW %>% 
                   filter(0 <= Time_Start & Time_Start < 20),num_time_steps=40, 
                          numcep = 12) %>% drop_na())$Humpback)


```




### Creating Training Data

```{r}
trainingdata = add_observations(whaledf = annotatedMelAW %>% 
                   filter(0 <= Time_Start & Time_Start < 200),num_time_steps=40, 
                          numcep = 12) %>% drop_na()

trainingdata = trainingdata %>% mutate(Humpback = as.factor(Humpback))

#trainingdata %>% filter(Humpback ==1)

table(trainingdata$Humpback)
```

### Creating Testing Data


```{r}

testingdata = add_observations(whaledf = annotatedMelAW %>% 
                   filter(200 <= Time_Start & Time_Start < 400),num_time_steps=40, 
                          numcep = 12) %>% drop_na()


testingdata = testingdata %>% mutate(Humpback = as.factor(Humpback))

table(testingdata$Humpback)

```


### Creating Cross-Validation Folds

```{r}
set.seed(326)
training_cvs = vfold_cv(trainingdata, v = 10)

```

## LDA

### Creating LDA Model

```{r}
#creating model
lda_mod1 = discrim_linear() %>%
  set_engine("MASS") %>%
  set_mode("classification")

#creating recipe
ldarecipe_1 = recipe(Humpback ~ ., data = trainingdata) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
ldaworkflow_1 = workflow() %>% 
  add_recipe(ldarecipe_1) %>% 
  add_model(lda_mod1)

#fitting model
ldafit_1 = ldaworkflow_1 %>% fit(trainingdata) 

#collinearity warning sometimes

#confusion matrix
trainingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, trainingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = ldapreds)
```


```{r}
trainingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, trainingdata)$.pred_class
  ) %>% select(Humpback, ldapreds, Time_Start, Time_End)


#ldaworkflow_1 %>% fit_resamples(training_cvs)


ldafitcv_1 = ldaworkflow_1 %>% fit_resamples(training_cvs,
                metrics = metric_set(accuracy, roc_auc, precision, specificity, recall)
                )
collect_metrics(ldafitcv_1)


2236/(2336 + 60)
```

### Predicting on Testing Data

```{r}

#confusion matrix
testingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, testingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = ldapreds)



testingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, testingdata)$.pred_class
  ) %>% 
  select(Humpback, ldapreds, Time_Start, Time_End)



testingdata %>%
  mutate(
    ldapreds = as.factor(predict(ldafit_1, testingdata)$.pred_class)
  ) %>% recall(truth = Humpback, estimate =  ldapreds, event_level = 'second')
```

### Computing Metrics on Validation Data Set

```{r}

```



### Create Output For Raven Visualization




```{r}
testingoutput = testingdata %>%
  mutate(ldapreds = predict(ldafit_1, testingdata)$.pred_class) %>% 
  dplyr::select(Time_Start, Time_End, Humpback, ldapreds)




#creating dummy low and high frequencies for now

testingoutput = testingoutput %>% mutate(low = rep(0, times = nrow(testingoutput)),
                         high = rep(1000, times = nrow(testingoutput)))



testingoutput = testingoutput %>% filter(ldapreds == 1) %>% 
  rename("Begin Time (s)" = Time_Start, "End Time (s)" = Time_End, 
         "Low Freq (Hz)" = low, "High Freq (Hz)" = high) %>% 
  dplyr::select(c("Begin Time (s)", "End Time (s)", 
                  "Low Freq (Hz)", "High Freq (Hz)"))



#create output 



write.table(testingoutput, "fulloverlapldapreds.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```

## KNN

### Creating KNN Model

```{r}
#creating model
knn_mod1 = nearest_neighbor(neighbors = 100) %>%
  set_engine("kknn") %>%
  set_mode("classification")

#creating recipe
knnrecipe_1 = recipe(Humpback ~ ., 
                     data = trainingdata) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
knnworkflow_1 = workflow() %>% 
  add_recipe(knnrecipe_1) %>% 
  add_model(knn_mod1)

#fitting model
knnfit_1 = knnworkflow_1 %>% fit(trainingdata) 


#confusion matrix
trainingdata %>%
  mutate(
    knnpreds = predict(knnfit_1, trainingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = knnpreds)




```


### KNN Predicting on Testing

```{r}

#confusion matrix
testingdata %>%
  mutate(
    knnpreds = predict(knnfit_1, testingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = knnpreds)

testingdata %>%
  mutate(
    knnpreds = as.factor(predict(knnfit_1, testingdata)$.pred_class)
  ) %>% recall(truth = Humpback, estimate =  knnpreds, event_level = 'second')


```

## Decision Tree



### Creating Model

```{r}
#model
decision_tree_mod1 = decision_tree() %>%
  set_engine("rpart") %>%
  set_mode("classification")

#recipe
treerecipe_1 = recipe(Humpback ~ ., 
                     data = trainingdata) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())

#workflow
tree_wflow = workflow() %>%
  add_model(decision_tree_mod1) %>% 
  add_recipe(treerecipe_1)


#fitting model
tree_fit_1 = tree_wflow %>% fit(trainingdata) 


#confusion matrix
trainingdata %>%
  mutate(
    treepreds = predict(tree_fit_1, trainingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = treepreds)
```

### DT Predicting on Testing



```{r}
#confusion matrix
testingdata %>%
  mutate(
    treepreds = predict(tree_fit_1, testingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = treepreds)

testingdata %>%
  mutate(
    treepreds = as.factor(predict(tree_fit_1, testingdata)$.pred_class)
  ) %>% recall(truth = Humpback, estimate =  treepreds, event_level = 'second')


```




