---
title: "ContextWindows"
format: html
editor: source
---

```{r}
library(tuneR)
library(seewave)
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
```

## Context Windows

The context window length, corresponds to a 3.9124-s long spectrogram image. A context window of 3.9124-s length is large enough to contain an entire unit but not large enough to contain two units (on average).

```{r}
AW = read_excel("C:/Users/lucas/Desktop/Thesis/Code/TrainingData/AWhb.xlsx")
AW

annotatedAW = read.csv("C:/Users/lucas/Desktop/Thesis/Code/annotatedmelAW.csv")

annotatedAW$Humpback = as.factor(annotatedAW$Humpback)

max(annotatedAW$instances)



```

```{r}
#randomize order of instances
max(annotatedAW$instances)  #392
392/10
set.seed(345)

shuffled_indices = sample(c(1:max(annotatedAW$instances)), 
                          size = max(annotatedAW$instances))


#split into cross validation folds
numfolds = 10
fold_assignments <- cut(shuffled_indices, breaks = numfolds, labels = FALSE)
table(fold_assignments)

#add them all to the cross validation fold
folds <- list()
for (i in 1:numfolds) {
  folds[[i]] <- shuffled_indices[fold_assignments == i]
}

folds


for (i in 1:k_folds) {
  test_indices <- folds[[i]]
  train_indices <- unlist(folds[-i]) # Combine all other folds

  train_data <- data[train_indices, ]
  test_data <- data[test_indices, ]

  # Train your model on train_data
  # Evaluate your model on test_data
  # Store performance metrics
  
  
}
```

## Manual RSET Object (Working Example)

```{r}
# Load the rsample package
library(rsample)
library(tibble) # rsample objects are tibbles

# 1. Define the data (using a simple data frame for demonstration)
df <- data.frame(x = c(1, 2, 3, 4, 5, 6),
                 y = c(0, 1, 1, 1, 0, 0))

df$y = as.factor(df$y)

# 2. Specify custom indices for the folds
# Fold 1: analysis = rows 1, 2, 3; assessment = rows 4, 5, 6
# Fold 2: analysis = rows 4, 5, 6; assessment = rows 1, 2, 3
indices <- list(
  list(analysis = c(1L, 2L, 3L), assessment = c(4L, 5L, 6L)),
  list(analysis = c(4L, 5L, 6L), assessment = c(1L, 2L, 3L))
)

# 3. Create the list of "rsplit" objects using make_splits()
splits <- lapply(indices, make_splits, data = df)

# 4. Create the manual rset object
custom_cv <- manual_rset(splits, ids = c("Fold 1", "Fold 2"))

# Print the resulting rset object
print(custom_cv)

#creating model
lda_mod1 = discrim_linear() %>%
  set_engine("MASS") %>%
  set_mode("classification")

#creating recipe
ldarecipe_1 = recipe(y ~ x, 
                     data = df) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
ldaworkflow_1 = workflow() %>% 
  add_recipe(ldarecipe_1) %>% 
  add_model(lda_mod1)

#fitting model
ldafit_1 = ldaworkflow_1 %>% fit(df) 


#confusion matrix
df %>%
  mutate(
    ldapreds = predict(ldafit_1, df)$.pred_class
  ) %>%
  conf_mat(truth = y,
           estimate = ldapreds)

```

## Cross Validating

```{r}
#fitting CVs
ldafitcv_1 = ldaworkflow_1 %>% fit_resamples(custom_cv,
                metrics = metric_set(accuracy, roc_auc, precision, specificity, recall)
                )

#results
collect_metrics(ldafitcv_1)
#0.643
```

## Manual RSET Object (Whales)

```{r}
# Load the rsample package
library(rsample)
library(tibble) # rsample objects are tibbles

# 1. Define the data (using a simple data frame for demonstration)
annotatedAW

# 2. Specify custom indices for the folds
# Fold 1: analysis = rows 1, 2, 3; assessment = rows 4, 5, 6
# Fold 2: analysis = rows 4, 5, 6; assessment = rows 1, 2, 3
indices <- list(
  list(analysis = c(1:2), assessment = c(3:6)),
  list(analysis = c(3:6), assessment = c(1:2))
)

indices
```

```{r}


# 3. Create the list of "rsplit" objects using make_splits()
splits <- lapply(indices, make_splits, data = df)

# 4. Create the manual rset object
custom_cv <- manual_rset(splits, ids = c("Fold 1", "Fold 2"))

# Print the resulting rset object
print(custom_cv)

#creating model
lda_mod1 = discrim_linear() %>%
  set_engine("MASS") %>%
  set_mode("classification")

#creating recipe
ldarecipe_1 = recipe(y ~ x, 
                     data = df) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
ldaworkflow_1 = workflow() %>% 
  add_recipe(ldarecipe_1) %>% 
  add_model(lda_mod1)

```

## Cross Validating

```{r}
#fitting CVs
ldafitcv_1 = ldaworkflow_1 %>% fit_resamples(custom_cv,
                metrics = metric_set(accuracy, roc_auc, precision, specificity, recall)
                )

#results
collect_metrics(ldafitcv_1)
#0.643
```

## Test AI Code for Converting to RSET Object

```{r}
# Load the rsample library
library(rsample)
library(tibble) # rset objects are tibbles

# 1. Define your original data
df <- data.frame(x = 1:6, y = letters[1:6])

# 2. Create your custom list of indices
# Each element in the list should be a list itself,
# containing 'analysis' (training) and 'assessment' (testing) indices.
# Indices should be integer vectors (e.g., 1L, 2L).
custom_indices <- list(
  list(analysis = c(1L, 2L, 4L, 5L), assessment = c(3L, 6L)),
  list(analysis = c(1L, 3L, 5L), assessment = c(2L, 4L, 6L))
)
class(custom_indices)

# 3. Use lapply and make_splits() to convert each index list to an rsplit object
# The 'data' argument tells make_splits which dataset the indices correspond to.
rsplits_list <- lapply(custom_indices, make_splits, data = df)
class(rsplits_list)

# 4. Use manual_rset() to combine the rsplit objects into an rset object
# You also need to provide character IDs for each split.
custom_rset <- manual_rset(
  splits = rsplits_list,
  ids = c("Split1", "Split2")
)

# Print the resulting rset object
print(custom_rset)
class(custom_rset)


rsplits_list <- lapply(folds, make_splits, data = annotatedAW)

manual_rset()

```

## Creating Wide Vector of Mel Annotations

```{r}

```

## Duration EDA

```{r}
ggplot(AW, aes(x = start_start_diff)) + geom_histogram(bins = 250)+ 
      scale_x_continuous(limits = c(0, 10))

```

```{r}
table(AW$overlap)
```

## Creating Non Overlapping Context Windows (3.9 seconds)

```{r}

```
