---
title: "BalancedDataSet"
format: html
editor: source
---



## Context Windows

The context window length, corresponds to a 3.9124-s long spectrogram image. A context window of 3.9124-s length is large enough to contain an entire unit but not large enough to contain two units (on average).

## Sourcing


```{r}
source(here("CreatingObservations.R"))
source(here("CreatingEvaluationWindows.R"))

```



## Loading Libraries

```{r}
library(tuneR)
library(seewave)
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)

library(tidyverse)
library(tidymodels)
library(kknn)
library(discrim)
```

## Setting Seed for Consistent Randomization

```{r}
set.seed(745)
```


## Loading In Data

```{r}
annotatedMelAW = read.csv(here::here("annotatedmelAWoverlap.csv"))
# first 30 minutes is in this data set

# annotatedMelAW %>% arrange(desc(Time_End))
```


## Creating Observations


```{r}


table((add_observations(whaledf = annotatedMelAW %>% 
                   filter(0 <= Time_Start & Time_Start < 20),num_time_steps=40, 
                          numcep = 12) %>% drop_na())$Humpback)


```




### Creating Balanced Training Data

```{r}
trainingdata = add_observations(whaledf = annotatedMelAW %>% 
                   filter(0 <= Time_Start & Time_Start < 200),num_time_steps=40, 
                          numcep = 12) %>% drop_na()

trainingdata = trainingdata %>% mutate(Humpback = as.factor(Humpback))

#trainingdata %>% filter(Humpback ==1)

table(trainingdata$Humpback)

numnegatives = nrow(trainingdata %>% filter(Humpback == 0))


# taking row numbers of positive observations

trainingpositives = trainingdata %>% filter(Humpback == 1)



sample(c(1:nrow(trainingpositives)), size = numnegatives)

trainingpositives[sample(c(1:nrow(trainingpositives)), size = numnegatives),]


trainingdatabalanced = rbind(trainingdata %>% filter(Humpback == 0),
                              trainingpositives[sample(c(1:nrow(trainingpositives)), size = numnegatives),]
                                  )

table(trainingdatabalanced$Humpback)
```

### Creating Testing Data


```{r}

testingdata = add_observations(whaledf = annotatedMelAW %>% 
                   filter(200 <= Time_Start & Time_Start < 400),num_time_steps=40, 
                          numcep = 12) %>% drop_na()


testingdata = testingdata %>% mutate(Humpback = as.factor(Humpback))

table(testingdata$Humpback)

```



### Creating Cross-Validation Folds

```{r}
training_cvs = vfold_cv(trainingdatabalanced, v = 10)

```

## LDA

### Creating LDA Model

```{r}
#creating model
lda_mod1 = discrim_linear() %>%
  set_engine("MASS") %>%
  set_mode("classification")

#creating recipe
ldarecipe_1 = recipe(Humpback ~ ., data = trainingdatabalanced) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
ldaworkflow_1 = workflow() %>% 
  add_recipe(ldarecipe_1) %>% 
  add_model(lda_mod1)

#fitting model
ldafit_1 = ldaworkflow_1 %>% fit(trainingdatabalanced) 

#collinearity warning sometimes

#confusion matrix
trainingdatabalanced %>%
  mutate(
    ldapreds = predict(ldafit_1, trainingdatabalanced)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = ldapreds)
```


```{r}
trainingdatabalanced %>%
  mutate(
    ldapreds = predict(ldafit_1, trainingdatabalanced)$.pred_class
  ) %>% dplyr::select(Humpback, ldapreds, Time_Start, Time_End)


#ldaworkflow_1 %>% fit_resamples(training_cvs)


ldafitcv_1 = ldaworkflow_1 %>% fit_resamples(training_cvs,
                metrics = metric_set(accuracy, roc_auc, precision, specificity, recall)
                )
collect_metrics(ldafitcv_1)


```


### Creating Evaluation Windows


```{r}
testingpredictions = testingdata %>%
  mutate(
    preds = predict(ldafit_1, testingdata)$.pred_class
  ) %>% dplyr::select(Time_Start, Time_End, Humpback, preds, Humpback)

```


```{r}


creating_eval_windows(testingpredictions, 1)

```


### Testing on Evaluation Windows

```{r}



evalwindows(testingpredictions, 1, 0.5)



as.numeric(nrow(testingpredictions %>% filter(200 <= Time_Start & Time_Start < 205) %>% 
  filter(Humpback == 1)) >= 1)

```


### Predicting on Testing Data

```{r}

#confusion matrix
testingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, testingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = ldapreds)



testingdata %>%
  mutate(
    ldapreds = predict(ldafit_1, testingdata)$.pred_class
  ) %>% 
  select(Humpback, ldapreds, Time_Start, Time_End)



testingdata %>%
  mutate(
    ldapreds = as.factor(predict(ldafit_1, testingdata)$.pred_class)
  ) %>% recall(truth = Humpback, estimate =  ldapreds, event_level = 'second')
```


### Create Output For Raven Visualization


```{r}
testingoutputlda = testingdata %>%
  mutate(ldapreds = predict(ldafit_1, testingdata)$.pred_class) %>% 
  dplyr::select(Time_Start, Time_End, Humpback, ldapreds)




#creating dummy low and high frequencies for now

testingoutputlda = testingoutputlda %>% mutate(low = rep(0, times = nrow(testingoutputlda)),
                         high = rep(1000, times = nrow(testingoutputlda)))



testingoutputlda = testingoutputlda %>% filter(ldapreds == 1) %>% 
  rename("Begin Time (s)" = Time_Start, "End Time (s)" = Time_End, 
         "Low Freq (Hz)" = low, "High Freq (Hz)" = high) %>% 
  dplyr::select(c("Begin Time (s)", "End Time (s)", 
                  "Low Freq (Hz)", "High Freq (Hz)"))



#create output 
write.table(testingoutputlda, "balancedfulloverlapldapreds.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```



## KNN

### Creating KNN Model

```{r}
#creating model
knn_mod1 = nearest_neighbor(neighbors = 100) %>%
  set_engine("kknn") %>%
  set_mode("classification")

#creating recipe
knnrecipe_1 = recipe(Humpback ~ ., 
                     data = trainingdatabalanced) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
knnworkflow_1 = workflow() %>% 
  add_recipe(knnrecipe_1) %>% 
  add_model(knn_mod1)

#fitting model
knnfit_1 = knnworkflow_1 %>% fit(trainingdatabalanced) 


#confusion matrix
trainingdatabalanced %>%
  mutate(
    knnpreds = predict(knnfit_1, trainingdatabalanced)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = knnpreds)




```


### KNN Predicting on Testing

```{r}

#confusion matrix
testingdata %>%
  mutate(
    knnpreds = predict(knnfit_1, testingdata)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = knnpreds)

testingdata %>%
  mutate(
    knnpreds = as.factor(predict(knnfit_1, testingdata)$.pred_class)
  ) %>% recall(truth = Humpback, estimate =  knnpreds, event_level = 'second')


```


### Create Output For Raven Visualization


```{r}
testingoutputknn = testingdata %>%
  mutate(knnpreds = predict(knnfit_1, testingdata)$.pred_class) %>% 
  dplyr::select(Time_Start, Time_End, Humpback, knnpreds)




#creating dummy low and high frequencies for now

testingoutputknn = testingoutputknn %>% mutate(low = rep(0, times = nrow(testingoutputknn)),
                         high = rep(1000, times = nrow(testingoutputknn)))



testingoutputknn = testingoutputknn %>% filter(knnpreds == 1) %>% 
  rename("Begin Time (s)" = Time_Start, "End Time (s)" = Time_End, 
         "Low Freq (Hz)" = low, "High Freq (Hz)" = high) %>% 
  dplyr::select(c("Begin Time (s)", "End Time (s)", 
                  "Low Freq (Hz)", "High Freq (Hz)"))



#create output 
write.table(testingoutputknn, "balancedfulloverlapknnpreds.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```
