---
title: "Model Tuning"
format: html
editor: source
---

## Loading Libraries

```{r}
library(tuneR)
library(seewave)
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)

library(tidyverse)
library(tidymodels)
library(kknn)
library(discrim)
```

## Setting Seed for Consistent Randomization

```{r}
set.seed(745)
```


## Loading In Data

```{r}
annotatedMelAW = read.csv(here::here("annotatedmelAWoverlap.csv"))
# first 30 minutes is in this data set

#annotatedMelAW %>% arrange(desc(Time_End))
```


## Creating Empty Data Frame


```{r}

#numcep is the number of mel coefficients used
#num_time_steps is the number of time steps recorded in an observation

#create_observations takes in the above parameters,
  #outputs a data frame with num_time_steps*numcep columns, 
  #and 'Time_Start', 'Time_End' and Humpback (0/1) columns

create_observations = function(numcep, num_time_steps){
  #testdf = data.frame(nrow = num_time_steps*numcep)
  #print(testdf)
  colnamelist = list()
  for (i in 1:numcep){
    for (j in 1:num_time_steps){
      #print(i)
      #print(j)
      #testdf = cbind(testdf, paste0("V1_",i))
      colnamelist = append(colnamelist, paste0("V", i, "_", j))
    }
  }
  #adding Time_Start and Time_End variables to end of each row
  colnamelist = append(colnamelist, "Time_Start")
  colnamelist = append(colnamelist, "Time_End")
  colnamelist = append(colnamelist, "Humpback")


  dummydf = data.frame(matrix(nrow = 1, ncol = num_time_steps*numcep + 3))
  
  colnames(dummydf) = colnamelist
  
  
  
  
  
  return(dummydf)
  
}
#emptydf = create_observations(numcep = 8, num_time_steps = 40)

#emptydf
```



## Creating Observations


```{r}

#function for specifically overlapped time steps (0.5 second increases, 0.5 second overlap)


# dummydf is the empty df created in the above 'create_observations' function
# whaledf is the full mel coefficient data frame
# num_time_steps is the number of time steps 
# numcep is the number of cepstrum coefficients



add_observations = function(whaledf, num_time_steps, numcep){
  
  
  dummydf = create_observations(numcep, num_time_steps)

  #getting initial times
  #filtering on initial times
  initials = whaledf %>% pull(Time_Start)
  
  # for each initial row in which an observation is desired
  for (i in initials){
    
    
    
    #i represents the start time of an initial observation (context window)
    
    #create observation by looping over the number of time steps
    
    #if statement to avoid including observations at end of the file
    if (nrow(whaledf %>% 
             filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps)) == 
        num_time_steps)
      {
        # determining whether the window has any whale call
        if (nrow(whaledf %>% 
             filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps) %>%
             filter(Humpback == 1)) >= 1){
          
        whale = 1
        } else {
        whale = 0
        }
   
          
      coeffs = whaledf %>% 
        filter(i <= Time_Start & Time_Start < i + 0.05*num_time_steps) %>% 
        dplyr::select(V1:paste0("V", numcep))
    
    
      #c(as.matrix()) creates a vector out of a data.frame, going down columns
      dummydf = rbind(dummydf, 
                      c(as.matrix(coeffs), i, i + 0.05*num_time_steps, whale))
      }
    
    
    #need to create conditionals for end of file
  }
  return(dummydf)
}


#positives = 
  
  



table((add_observations(whaledf = annotatedMelAW %>% 
                   filter(0 <= Time_Start & Time_Start < 200),num_time_steps=40, 
                          numcep = 12) %>% drop_na())$Humpback)


```




### Creating Balanced Training Data

```{r}
trainingdata = add_observations(whaledf = annotatedMelAW %>% 
                   filter(0 <= Time_Start & Time_Start < 200),num_time_steps=40, 
                          numcep = 12) %>% drop_na()

trainingdata = trainingdata %>% mutate(Humpback = as.factor(Humpback))


table(trainingdata$Humpback)

numnegatives = nrow(trainingdata %>% filter(Humpback == 0))


# taking row numbers of positive observations

trainingpositives = trainingdata %>% filter(Humpback == 1)



# choosing random sample of positive observations
trainingpositives[sample(c(1:nrow(trainingpositives)), size = numnegatives),]


trainingdatabalanced = rbind(trainingdata %>% filter(Humpback == 0),
                              trainingpositives[sample(c(1:nrow(trainingpositives)), size = numnegatives),]
                                  )

table(trainingdatabalanced$Humpback)
```

### Creating Testing Data


```{r}

testingdata = add_observations(whaledf = annotatedMelAW %>% 
                   filter(200 <= Time_Start & Time_Start < 400),num_time_steps=40, 
                          numcep = 12) %>% drop_na()


testingdata = testingdata %>% mutate(Humpback = as.factor(Humpback))

table(testingdata$Humpback)

```



### Creating Training Data Cross-Validation Folds

```{r}
training_cvs = vfold_cv(trainingdatabalanced, v = 10)

```



### Creating Evaluation Windows on Testing Data





```{r}
creating_eval_windows = function(predictions, length){
  # initial times
  initial = predictions %>% pull(Time_Start) %>% first()
  final = predictions %>% pull(Time_Start) %>% last()
  initials = c()
  starttime = initial
  
  # for all initial times that fit in the testing window:
  while (starttime + length <= final){
  
    newtime = starttime
    initials = c(initials, newtime)
    starttime = starttime + length
    
  }
  return(initials)
  
}

#creating_eval_windows(testingpredictions, 1)

```


## Creating Rolling Aggregation Windows


```{r}

# predictions is a matrix with time start, time end, prediction, and Humpback
# length is the length of each evaluation window
# threshold is the proportion of positives required for the threshold window


testing_eval_windows = function(predictions, length, threshold){
  
  initials = creating_eval_windows(predictions, length)
  
  init_matrix = matrix(nrow = 0, ncol = 5)
  
  #create data frame with state time, end time, proportion, and 0/1 based on prop vs threshold, and truth

  results = data.frame(init_matrix)
  
  for (i in initials){
    #print(predictions %>% filter(i <= Time_Start & Time_Start < i + length))
    
    # if enough windows predicted a whale in this evaluation window
    if (predictions %>% filter(i <= Time_Start & Time_Start < i + length) %>% 
        summarise(proportion = sum(preds == 1) / n()) %>% pull(proportion) > threshold){
      

      row = c("Time_Start" = i, 
              "Time_End" = i + length, 
              "proportion" = predictions %>% filter(i <= Time_Start & Time_Start < i + length) %>% 
        summarise(proportion = sum(preds == 1) / n()) %>% pull(proportion), 
              "prediction" = 1,
              "Humpback" = as.numeric(nrow(predictions %>% filter(i <= Time_Start & Time_Start < i + length) %>% 
  filter(Humpback == 1)) >= 1))
    
    # if not enough windows predicted a whale in this evaluation window
  
    } else{
      row = c("Time_Start" = i, 
        "Time_End" = i + length, 
        "proportion" = predictions %>% filter(i <= Time_Start & Time_Start < i + length) %>% 
        summarise(proportion = sum(preds == 1) / n()) %>% pull(proportion), 
        "prediction" = 0,
        "Humpback" = as.numeric(nrow(predictions %>% filter(i <= Time_Start & Time_Start < i + length) %>% 
  filter(Humpback == 1)) >= 1))
      
    }
    #print(row)
    results = rbind(results, row)
  }
  colnames(results) = c("Time_Start", "Time_End", "proportion", "prediction", "Humpback")
  results = results%>% mutate(prediction = as.factor(prediction),
                              Humpback = as.factor(Humpback))
  return(results)
  

  
  
  # end observation at start + length
  # if number of time steps that predicted a whale is >= proportion
    # make this observation a 1
    # else: make it 0
  # start new observation at next time step
  # repeat until end of data file
  
  
  
  
  
}

```





## Creating Random Forest Model


```{r}

#creating model
rfmod_1 <- rand_forest(trees = 1000) %>%
  set_engine("ranger") %>%               
  set_mode("classification")            

#creating recipe
rfrecipe_1 <- recipe(Humpback ~ ., data = trainingdatabalanced) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
rfworkflow_1 <- workflow() %>%
  add_model(rfmod_1) %>%
  add_recipe(rfrecipe_1)



#fitting model
rffit_1 <- rfworkflow_1 %>% fit(data = trainingdatabalanced)


```



### Cross Validating

```{r}
trainingdatabalanced %>%
  mutate(
    rfpreds = predict(rffit_1, trainingdatabalanced)$.pred_class
  ) %>% dplyr::select(Humpback, rfpreds, Time_Start, Time_End)


#ldaworkflow_1 %>% fit_resamples(training_cvs)


rffitcv_1 = rfworkflow_1 %>% fit_resamples(training_cvs,
                metrics = metric_set(accuracy, roc_auc, precision, specificity, recall)
                )
collect_metrics(rffitcv_1)


```



### Testing Using Evaluation Windows


```{r}


rfwindowpredictions = testing_eval_windows(testingpredictionsrf, length = 1, threshold = 0.3)

# confusion matrix
rfwindowpredictions %>%
  conf_mat(truth = Humpback,
           estimate = prediction)

# recall metric
rfwindowpredictions %>% 
  recall(truth = Humpback, estimate = prediction, event_level = 'second')
```

