---
title: "Rolling Eval Windows"
format: html
editor: source
---



This file takes in an object of predictions, creates discrete(?) windows to create a more accurate prediction data set.


## Loading Libraries

```{r}
library(tuneR)
library(seewave)
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
library(ranger)
library(tidyverse)
library(tidymodels)
library(kknn)
library(discrim)
```

## Sourcing

```{r}
source(here::here("ContextWindows", "ContextWindowFunctions.R"))
source(here::here("EvaluationWindowFunctions.R"))

```



## Setting Seed for Consistent Randomization

```{r}
set.seed(288)
```


## Loading In Data

```{r}
#annotatedMelAW = read.csv(here::here("annotatedmelAWoverlap.csv"))

data = read.csv(here::here("60minutes.csv"))

#annotatedMelAW = read.csv(here::here("annotatedmelAWlowerhzoverlap.csv"))
# first 30 minutes is in this data set

#annotatedMelAW %>% arrange(desc(Time_End))
```




### Creating Balanced Training Data

```{r}
# start and end time are in seconds

trainingdata = add_observations(whaledf = data,
                                start_time = 0,
                                end_time = 120,
                                num_time_steps=40, 
                                numcep = 12,
                                step_size = 0.05)

table(trainingdata$Humpback)

numnegatives = nrow(trainingdata %>% filter(Humpback == 0))


# taking row numbers of positive observations

trainingpositives = trainingdata %>% filter(Humpback == 1)


sample(c(1:nrow(trainingpositives)), size = numnegatives)

trainingpositives[sample(c(1:nrow(trainingpositives)), size = numnegatives),]


trainingdatabalanced = rbind(trainingdata %>% filter(Humpback == 0),
                              trainingpositives[sample(c(1:nrow(trainingpositives)), size = numnegatives),]
                                  )

table(trainingdatabalanced$Humpback)
```

### Creating Testing Data


```{r}
testingdata = add_observations(whaledf = data,
                                start_time = 300,
                                end_time = 360,
                                num_time_steps=40, 
                                numcep = 12,
                                step_size = 0.05)

table(testingdata$Humpback)

```

### Creating Cross-Validation Folds

```{r}
training_cvs = vfold_cv(trainingdatabalanced, v = 10)

```


## LDA

### Creating LDA Model

```{r}
#creating model
lda_mod1 = discrim_linear() %>%
  set_engine("MASS") %>%
  set_mode("classification")

#creating recipe
ldarecipe_1 = recipe(Humpback ~ ., data = trainingdatabalanced) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
ldaworkflow_1 = workflow() %>% 
  add_recipe(ldarecipe_1) %>% 
  add_model(lda_mod1)

#fitting model
ldafit_1 = ldaworkflow_1 %>% fit(trainingdatabalanced) 

#collinearity warning sometimes

#confusion matrix
trainingdatabalanced %>%
  mutate(
    ldapreds = predict(ldafit_1, trainingdatabalanced)$.pred_class
  ) %>%
  conf_mat(truth = Humpback,
           estimate = ldapreds)
```

### Cross Validating

```{r}
trainingdatabalanced %>%
  mutate(
    ldapreds = predict(ldafit_1, trainingdatabalanced)$.pred_class
  ) %>% dplyr::select(Humpback, ldapreds, Time_Start, Time_End)


#ldaworkflow_1 %>% fit_resamples(training_cvs)


ldafitcv_1 = ldaworkflow_1 %>% fit_resamples(training_cvs,
                metrics = metric_set(accuracy, roc_auc, precision, specificity, recall)
                )
collect_metrics(ldafitcv_1)


```


### Predicting on Testing Data

```{r}

ldapredictions = testingdata %>%
  mutate(
    preds = predict(ldafit_1, testingdata)$.pred_class
  ) %>% dplyr::select(Time_Start, Time_End, preds, Humpback)


ldawindowpredictions = testing_eval_windows(ldapredictions, length = 1, threshold = 0.3)

source(here::here("CustomDenominator.R"))

manual_annotations = read_excel(here::here("TrainingData", "AW", "AWhb.xlsx"))

custom_denominator(manual_annotations, ldapredictions,
                   pred_start_time = 30, pred_end_time = 58, threshold = 0.5)

ldawindowpredictions %>%
  conf_mat(truth = Humpback,
           estimate = prediction)

ldawindowpredictions %>% 
  recall(truth = Humpback, estimate = prediction, event_level = 'second')
```

### Create Output For Raven Visualization


```{r}
ldatestingoutput = ldawindowpredictions %>% 
  dplyr::select(Time_Start, Time_End, Humpback, prediction)




#creating dummy low and high frequencies for now

ldatestingoutput = ldatestingoutput %>% mutate(low = rep(0, times = nrow(ldatestingoutput)),
                         high = rep(2000, times = nrow(ldatestingoutput)))



ldatestingoutput = ldatestingoutput %>% filter(prediction == 1) %>% 
  rename("Begin Time (s)" = Time_Start, "End Time (s)" = Time_End, 
         "Low Freq (Hz)" = low, "High Freq (Hz)" = high) %>% 
  dplyr::select(c("Begin Time (s)", "End Time (s)", 
                  "Low Freq (Hz)", "High Freq (Hz)"))



#create output 
write.table(ldatestingoutput, "ldaevalwindowpreds.txt", sep = "\t", row.names = FALSE, quote = FALSE)
```
