---
title: "Random Forest Model"
format: html
editor: source
---



## Loading Libraries

```{r}
library(tuneR)
library(seewave)
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
library(ranger)
library(tidyverse)
library(tidymodels)
library(kknn)
library(discrim)
```

## Sourcing

```{r}
source(here::here("ContextWindows", "ContextWindowFunctions.R"))
source(here::here("EvaluationWindowFunctions.R"))
source(here::here("CustomDenominator.R"))
```



## Setting Seed for Consistent Randomization

```{r}
set.seed(656)
```


## Loading In Data

```{r}
#annotatedMelAW = read.csv(here::here("annotatedmelAWoverlap.csv"))

# data = read.csv(here::here("60minutes.csv"))

data = read.csv(here::here("FullMelAW.csv"))

#annotatedMelAW = read.csv(here::here("annotatedmelAWlowerhzoverlap.csv"))
# first 30 minutes is in this data set

#annotatedMelAW %>% arrange(desc(Time_End))
```




### Creating Balanced Training Data

```{r}
# start and end time are in seconds

trainingdata = add_observations(whaledf = data,
                                start_time = 0,
                                end_time = 700,
                                num_time_steps=20, 
                                numcep = 12,
                                step_size = 0.05)

table(trainingdata$Humpback)

numnegatives = nrow(trainingdata %>% filter(Humpback == 0))


# taking row numbers of positive observations

trainingpositives = trainingdata %>% filter(Humpback == 1)


sample(c(1:nrow(trainingpositives)), size = numnegatives)

trainingpositives[sample(c(1:nrow(trainingpositives)), size = numnegatives),]


trainingdatabalanced = rbind(trainingdata %>% filter(Humpback == 0),
                              trainingpositives[sample(c(1:nrow(trainingpositives)), size = numnegatives),]
                                  )

table(trainingdatabalanced$Humpback)
```

### Creating Testing Data


```{r}
testingdata = add_observations(whaledf = data,
                                start_time = 2000,
                                end_time = 2500,
                                num_time_steps=20, 
                                numcep = 12,
                                step_size = 0.05)

table(testingdata$Humpback)

```

### Creating Cross-Validation Folds

```{r}
training_cvs = vfold_cv(trainingdatabalanced, v = 10)

```





## Random Forest

### Creating RF Model

```{r}
library(randomForest)

#creating model
rfmod_1 <- rand_forest(min_n = 100, trees = 2000,
                       mtry = floor(sqrt(ncol(trainingdatabalanced) - 3))) %>%
  set_engine("ranger") %>%               
  set_mode("classification")            

#creating recipe
rfrecipe_1 <- recipe(Humpback ~ ., data = trainingdatabalanced) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
rfworkflow_1 <- workflow() %>%
  add_model(rfmod_1) %>%
  add_recipe(rfrecipe_1)



#fitting model
rffit_1 <- rfworkflow_1 %>% fit(data = trainingdatabalanced)


rffit_1 |>
  extract_fit_engine()

print(rfmod_1)
```



### Cross Validating

```{r}
trainingdatabalanced %>%
  mutate(
    rfpreds = predict(rffit_1, trainingdatabalanced)$.pred_class
  ) %>% dplyr::select(Humpback, rfpreds, Time_Start, Time_End)



rffitcv_1 = rfworkflow_1 %>% fit_resamples(training_cvs,
                metrics = metric_set(accuracy, roc_auc, precision, specificity, recall)
                )
collect_metrics(rffitcv_1)


```





### Predicting on Testing Data Using Evaluation Windows

```{r}

rfpredictions = testingdata %>%
  mutate(
    preds = predict(rffit_1, testingdata)$.pred_class
  ) %>% dplyr::select(Time_Start, Time_End, preds, Humpback)


rfwindowpredictions = testing_eval_windows(rfpredictions, length = 1, threshold = 0.5)

rfwindowpredictions

rfwindowpredictions %>%
  conf_mat(truth = Humpback,
           estimate = prediction)

rfwindowpredictions %>% 
  recall(truth = Humpback, estimate = prediction, event_level = 'second')
```


### Predicting on Individual Whale Calls


```{r}
manual_annotations = read_excel(here::here("TrainingData", "AW", "AWhb.xlsx"))

custom_denominator(manual_annotations, rfpredictions,
                   pred_start_time = min(rfpredictions$Time_Start), 
                   pred_end_time = max(rfpredictions$Time_End), 
                   threshold = 0.5)
```



### Create Output For Raven Visualization

```{r}

raven_output = function(testingoutput, start, end){

  rftestingoutput = rfwindowpredictions %>% 
    dplyr::select(Time_Start, Time_End, Humpback, prediction)
  
  
  #creating dummy low and high frequencies
  
  rftestingoutput = rftestingoutput %>% mutate(low = rep(0, times = nrow(rftestingoutput)),
                           high = rep(2000, times = nrow(rftestingoutput)))
  
  
  rftestingoutput = rftestingoutput %>% filter(prediction == 1) %>% 
    rename("Begin Time (s)" = Time_Start, "End Time (s)" = Time_End, 
           "Low Freq (Hz)" = low, "High Freq (Hz)" = high) %>% 
    dplyr::select(c("Begin Time (s)", "End Time (s)", 
                    "Low Freq (Hz)", "High Freq (Hz)"))
  
  
  #create output 
  write.table(rftestingoutput, paste0("rfevalwindowpreds", start, end, ".txt"), 
              sep = "\t", row.names = FALSE, quote = FALSE)
}

raven_output(testingoutput = rftestingoutput,
             start = min(rfwindowpredictions$Time_Start),
             end = max(rfwindowpredictions$Time_End))

```



```{r}
test_func = function(input){
  print(paste0(input, "dflkgjhd"))
  
  
  
  
}

test_func(4598)
```

