---
title: "Random Forest Model Tuning"
format: html
editor: source
---


- function for specifying context window length (num_time_steps), aggregation length, and threshold link

- then compare recall, F2, and how many whales are correctly predicted

- start with 2 each, create data frame of all combinations and pass it through the function



## Loading Libraries

```{r}
library(tuneR)
library(seewave)
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
library(ranger)
library(tidyverse)
library(tidymodels)
library(kknn)
library(discrim)
```

## Sourcing

```{r}
source(here::here("ContextWindows", "ContextWindowFunctions.R"))
source(here::here("EvaluationWindowFunctions.R"))
source(here::here("CustomDenominator.R"))
```



## Setting Seed for Consistent Randomization

```{r}
set.seed(716)
```




## Loading In Data

```{r}
data = read.csv(here::here("annotatedmelAW.csv"))

# data = read.csv(here::here("60minutes.csv"))

# data = read.csv(here::here("FullMelAW.csv"))

#annotatedMelAW = read.csv(here::here("annotatedmelAWlowerhzoverlap.csv"))
# first 30 minutes is in this data set

#annotatedMelAW %>% arrange(desc(Time_End))
```


## Choosing 


## Function


```{r}
trainingdata = add_observations(whaledf = data,
                                start_time = 0,
                                end_time = 200,
                                num_time_steps='specify', 
                                numcep = 12,
                                step_size = 0.05)
```


## ...
