---
title: "Random Forest"
format: html
editor: source
---


## Loading Libraries

```{r}
library(tuneR)
library(seewave)
library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
library(ranger)
library(tidyverse)
library(tidymodels)
library(discrim)
library(randomForest)
```


## Setting Seed for Consistent Randomization


```{r}
set.seed(300)
```


## Sourcing

```{r}
source(here::here("ContextWindows", "ContextWindowFunction.R"))
source(here::here("EvaluationWindows", "RollingAverage.R"))
source(here::here("TestingPredictionIntervals.R"))
```


## Loading In Data

```{r}
data_train = read.csv(here::here("GrayWhaleMethod", 
                                 "TrainingSet", 
                                 "GrayWhaleTrainingData.csv")) %>% 
  mutate(Humpback = as.factor(Humpback))


data_test = read.csv(here::here("GrayWhaleMethod", 
                                "EvaluationSet",
                                "GrayWhaleEvaluationData.csv")) %>%
  mutate(Humpback = as.factor(Humpback))


#annotatedMelAW %>% arrange(desc(Time_End))
```




## Creating Training Data

```{r}

trainingdata = add_observations_linear_tolerance(whaledf = data_train,
                                start_time = 0,
                                end_time =  7199.9,
                                num_time_steps = 1, 
                                numcep = 13,
                                step_size = 0.1)




table(trainingdata$Humpback)

```



## Creating Testing Data


```{r}
testingdata = add_observations_linear_tolerance(whaledf = data_test,
                                start_time = 7200,
                                end_time =  max(data_test$Time_End),
                                num_time_steps = 1, 
                                numcep = 13,
                                step_size = 0.1)

testingdata

table(testingdata$Humpback)
```


## Loading in Annotations

```{r}
manual_annotations_evaluation = 
  read_excel(here::here("OrigData", "AW", "AWhb.xlsx")) %>% 
    rename(Begin_Time = 'Begin Time (s)', End_Time = 'End Time (s)') %>%
  filter(Begin_Time >= min(testingdata$Time_Start) & End_Time <= max(testingdata$Time_End))
```


## Creating Final Model

```{r}

#creating model
rfmod_1 <- rand_forest(min_n = 50, trees = 1250,
                       mtry = floor(sqrt(ncol(trainingdata) - 3))) %>%
  set_engine("ranger") %>%               
  set_mode("classification")            

#creating recipe
rfrecipe_1 <- recipe(Humpback ~ ., data = trainingdata) %>%
  step_rm(Time_Start, Time_End) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_dummy(all_factor_predictors())


#creating workflow
rfworkflow_1 <- workflow() %>%
  add_model(rfmod_1) %>%
  add_recipe(rfrecipe_1)



#fitting model
rffit_1 <- rfworkflow_1 %>% fit(data = trainingdata)


```







## Predicting on Testing Data

```{r}
predictionsdf = testingdata %>% mutate(preds = predict(rffit_1, testingdata)$.pred_class)
table(predictionsdf$preds)



predictionsdf %>%
  conf_mat(truth = Humpback, estimate = preds)
```


##Metrics Type 1

```{r}

data.frame(recall = recall(predictionsdf, truth = Humpback, estimate = preds, event_level = "second")$.estimate,
           precision = precision(predictionsdf, truth = Humpback, estimate = preds, event_level = "second")$.estimate, 
           F1 = f_meas(predictionsdf, truth = Humpback, estimate = preds, beta = 1, event_level = "second")$.estimate, 
           F2 = f_meas(predictionsdf, truth = Humpback, estimate = preds, beta = 2, event_level = "second")$.estimate
)
```







## Metrics Type 2


```{r}
whale_call_metrics_type2 <- function(predictionsdf) {
  # agg_window_length <- as.numeric(row[1])
  # agg_window_threshold <- as.numeric(row[2])

  # rolling_average <- rolling_avg_df(
  #   predictionsdf,
  #   agg_window_length,
  #   agg_window_threshold
  # )
  # 
  # print(rolling_average)

  # prediction_windows <- creating_prediction_windows(
  #   rolling_average,
  #   time_start = Time_Start,
  #   time_end = Time_End,
  #   value_col = rolling_avg_prediction
  # )
  
  
  positive_windows = predictionsdf %>% filter(preds == 1)
  
  print(positive_windows)

  # conf_matrix_table <- confusion_from_intervals(
  #   data_true = manual_annotations_evaluation,
  #   data_pred = prediction_windows,
  #   true_interval_start = Begin_Time,
  #   true_interval_end = End_Time,
  #   pred_interval_start = start_time,
  #   pred_interval_end = end_time
  # )
  # 

  conf_matrix_table <- confusion_from_intervals(
  data_true = manual_annotations_evaluation,
  data_pred = positive_windows,
  true_interval_start = Begin_Time,
  true_interval_end = End_Time,
  pred_interval_start = Time_Start,
  pred_interval_end = Time_End
  )
  

  TP <- conf_matrix_table %>% dplyr::pull(TP)
  FN <- conf_matrix_table %>% dplyr::pull(FN)
  FP <- conf_matrix_table %>% dplyr::pull(FP)

  print(TP)
  print(FN)
  print(FP)
  
  num_pred_windows <- nrow(positive_windows)
  num_calls_detected = TP

  if (TP + FN == 0) {
    recall <- NA_real_
  } else {
    recall <- TP / (TP + FN)
  }

  if (TP + FP == 0) {
    precision <- NA_real_
  } else {
    precision <- TP / (TP + FP)
  }

  if (is.na(recall) || is.na(precision) || (recall + precision == 0)) {
    F1 <- NA_real_
  } else {
    F1 <- 2 * recall * precision / (recall + precision)
  }

  if (is.na(recall) || is.na(precision) || (4 * precision + recall == 0)) {
    F2 <- NA_real_
  } else {
    F2 <- 5 * precision * recall / (4 * precision + recall)
  }

  data.frame(
    TP = TP,
    FN = FN,
    FP = FP,
    num_pred_windows = num_pred_windows,
    recall = recall,
    precision = precision,
    F1 = F1
  )
}

whale_call_metrics_type2(predictionsdf)


```



## Examining How Long to Listen


```{r}

windows_temp = creating_prediction_windows(
  predictionsdf,
  time_start = Time_Start,
  time_end = Time_End,
  value_col = preds
)


windows_temp = windows_temp %>% mutate(time_length = end_time - start_time)


sum(windows_temp$time_length)
```

## Calculating How Much Downtime Successfully Predicted


```{r}
predictionsdf %>% filter(preds == 0, Humpback == 0)


true_negative_windows_temp = creating_TN_prediction_windows(
  predictionsdf,
  time_start = Time_Start,
  time_end = Time_End,
  value_col = preds,
  value_col2 = Humpback
)

true_negative_windows_temp = true_negative_windows_temp %>% 
  mutate(time_length = end_time - start_time)

sum(true_negative_windows_temp$time_length)

```




## Saving Predictions of Best Parameters


```{r}

prediction_windows_best <- creating_prediction_windows(
  predictionsdf,
  time_start = Time_Start,
  time_end = Time_End,
  value_col = preds
)

predictionsdf %>% select(Time_Start, Time_End, Humpback, preds) %>% filter(preds == 1)

prediction_windows_best
```




## Create Output For Raven Visualization

```{r}

raven_output = function(windows){
  print(windows)
  #creating dummy low and high frequencies
  windows = windows %>% 
    rename("Begin Time (s)" = start_time, "End Time (s)" = end_time) %>% 
    mutate("Low Freq (Hz)" = 0, "High Freq (Hz)" = 2000) %>% 
    dplyr::select(c("Begin Time (s)", "End Time (s)", 
                    "Low Freq (Hz)", "High Freq (Hz)"))
  
  
  #create output 
  write.table(windows, 
              paste0("graywhalemethodrfevalwindowpreds", 
                     "len", agg_window_length_best, 
                     "thres", agg_window_threshold_best,
                     ".txt"), 
              sep = "\t", row.names = FALSE, quote = FALSE)
}

raven_output(windows = prediction_windows_best)

```